---
title: "Lab 03 - Shapefiles week 1"
author: "Francisco SAntamarina"
date: "February 16, 2017"
output: github_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load the necessary packages and dataset for questions 1 and 2.
```{r, warning=F, message = F }
#setwd( "//hd.ad.syr.edu/02/7796a1/Documents/Desktop/DDM/DDM II" )
library( maptools )
library( sp )

#dir.create( "shapefiles" )
#setwd( "~/Desktop/DDM/DDM II/shapefiles.old" )

```

### Question 1: What are the demographics of neighborhoods in Syracuse?

#### Download a TIGER shapefile of census tracts in Syracuse (Onondaga County)
```{r}
#download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2010/TRACT/2010/tl_2010_36067_tract10.zip", "onondaga census tracts.zip" )
#unzip( "onondaga census tracts.zip" )
#file.remove( "onondaga census tracts.zip" )

dir()

syr <- readShapePoly( fn="tl_2010_36067_tract10", 
         proj4string=CRS("+proj=longlat +datum=WGS84") )
#syr <- readShapePoly( fn="01-05-2015", 
#         proj4string=CRS("+proj=longlat +datum=WGS84") )

plot( syr,  border="gray80" )
```

#### Download three demographic variables of interest, your choice, using the censusapi package
```{r}
# Francisco Santamarina's unique census key
censuskey <- "b0387a9f8213a8adc3e0f940b217cc73f9f50a6f"

# Variables to get: Households that received Food Stamps/SNAP in the past 12 months, Median Age, Median Family Income in the past 12 months, Aggregate travel time to work (in minutes), Public transportation to work (in minutes)
#myvars <- c( "NAME",  "B22003_002E", "B01002A_001E", "B19113_001E", "B08136_001E", "B08136_007E"  )

# Variables to get: Household received Food Stamps/SNAP in the past 12 months, Median Age, Median Family Income in the past 12 months
myvars <- c( "NAME",  "B22003_002E", "B01002A_001E", "B19113_001E" )


data2015 <- getCensus(name="acs5", vintage=2015, key=censuskey, 
    vars=myvars, 
    region="tract:*", # "*" means all available, so this will get all available tracts
    regionin="state:36" #  "36" is the FIPS code for the state of New York
    )
head( data2015 )
colnames( data2015 ) <- c( "Name", "State", "County", "Tract", "SNAP", "Median.Age", "Median.Family.Income" )

# Create a subset of data for Onondaga County
onondaga2015 <- data2015[ data2015$County == "067", ] #067 is the Onondaga County Code, within New York
head( onondaga2015 )

# Variables:
## Household received SNAP
## Median Age
## Median Family Income
#### Later: Travel time to work (in minutes)
#### Later: Public transportation to work (in minutes)
#### Later: Income < poverty level last 12 months
#### Later: Income > poverty level last 12 months
```

#### Create choropleth maps that include a title and legend for each variable

```{r}
map.choices <- c( "Quantile", "Standard Deviation", "Equal Interval")
```


Households that received Food Stamps/SNAP in the past 12 months
```{r}
min( onondaga2015$SNAP, na.rm = TRUE  )
max( onondaga2015$SNAP, na.rm = TRUE  )

# Assign colors: brightest green is least SNAP households, darkest green is most SNAP households
color.function.SNAP <- colorRampPalette( c( "chartreuse", "darkgreen" )  )
col.ramp.SNAP <- color.function.SNAP( 5 ) # Number of groups is 5

# Break up the values of the vectors into 5 levels according to their ranks and apply to each level one of the 5 colors.
color.SNAP <- cut( onondaga2015$SNAP , breaks = 5, labels = col.ramp.SNAP ) 
color.SNAP <- as.character( color.SNAP ) # Convert the assigned colors to each value into character

# Match the various assigned level values to the proper tracts
# SNAP.ordered <- match(   )   Skip this step for TIGER files?

plot( syr,  border = "gray80", col=color.SNAP )
title( main = "Households that received Food Stamps/SNAP in the past 12 months" )
map.scale( metric = F, ratio = T, relwidth = 0.15, cex = 0.5 )
legend.text.SN = c( "-0.7 - 140", "141 - 280", "280 - 421", "421 - 561", "561 - 702" )
legend( "bottomright",
        pch = 19, pt.cex = 1.5, cex = 0.7,
        bg = "blanchedalmond",
        legend = legend.text.SN,
        col = col.ramp.SNAP,
        box.col = "blanchedalmond",
        title = "Food Stamp Amounts"
)
```


Median Age
```{r}
min( onondaga2015$Median.Age, na.rm = TRUE  )
max( onondaga2015$Median.Age, na.rm = TRUE  )

color.function.MA <- colorRampPalette( c( "lightsalmon", "mediumorchid4" )  )
col.ramp.MA <- color.function.MA( 5 ) # Number of groups is 5

# Break up the values of the vectors into 5 levels according to their ranks and apply to each level one of the 5 colors.
color.MA <- cut( onondaga2015$Median.Age , breaks = 5, labels = col.ramp.MA ) 
color.MA <- as.character( color.MA ) # Convert the assigned colors to each value into character

# Match the various assigned level values to the proper tracts
# MA.ordered <- match(   )   Skip this step for TIGER files?

plot( syr,  border = "gray80", col=color.MA )
title( main = "Median Ages of Census Tracts" )
map.scale( metric = F, ratio = T, relwidth = 0.15, cex = 0.5 )
legend.text.MA = c( "19.5 - 28.7", "28.8 - 37.9", "37.9 - 47.2", "47.2 - 56.4", "56.4 - 65.6" )
legend( "bottomright",
        pch = 19, pt.cex = 1.5, cex = 0.7,
        bg = "blanchedalmond",
        legend = legend.text.MA,
        col = col.ramp.MA,
        box.col = "blanchedalmond",
        title = "Median Age Groups"
)
```


Median Family Income in the past 12 months
```{r}
min( onondaga2015$Median.Family.Income, na.rm = TRUE  )
max( onondaga2015$Median.Family.Income, na.rm = TRUE  )

color.function.MFI <- colorRampPalette( c( "cyan", "blue4" )  )
col.ramp.MFI <- color.function.MFI( 5 ) # Number of groups is 5

# Break up the values of the vectors into 5 levels according to their ranks and apply to each level one of the 5 colors.
color.MFI <- cut( onondaga2015$Median.Family.Income , breaks = 5, labels = col.ramp.MFI ) 
color.MFI <- as.character( color.MFI ) # Convert the assigned colors to each value into character
mfiTest <- cut( 
                onondaga2015$Median.Family.Income , 
                breaks = 5, 
                dig.lab = 9 #determines the number of digits to show when no labels
              )
levels(mfiTest)

# Match the various assigned level values to the proper tracts
# MFI.ordered <- match(   )   Skip this step for TIGER files?

plot( syr,  border = "gray80", col=color.MFI )
title( main = "Median Family Income in the Past 12 Months" )
map.scale( metric = F, ratio = T, relwidth = 0.15, cex = 0.5 )
legend.text.MFI = c( "$14,980 - $42,249", "$42,250 - $69,382", "69,383 - 96,515", "96,516 - 123,648", "123,649 - 150,916" )
legend( "bottomright",
        pch = 19, pt.cex = 1.5, cex = 0.7,
        bg = "blanchedalmond",
        legend = legend.text.MFI,
        col = col.ramp.MFI,
        box.col = "blanchedalmond",
        title = "Median Family\nIncome Groups"
)
```

