---
title: "Lab 06 - Creating Buffers"
author: "Francisco Santamarina"
date: "March 09, 2017"
output:
  html_document:
    df_print: paged
    keep_md: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F )
```

## Part 1

Load the necessary packages.
```{r}
#library( censusAPI)
#library( ggmap )
#library( jsonlite )
library( maps )
library( maptools )
library( measurements )
#library( memisc )
#library( RCurl )
library( raster )
library( rgdal )
library( rgeos )
library( sp )
#library( spatialEco ) 

```

Load and clean the necessary dataset.
```{r}
# Set the working directory
#setwd( "//hd.ad.syr.edu/02/7796a1/Documents/Desktop/DDM/DDM II/Lab 06" )
setwd("~/Graduate School/PAI 690 Independent Study_DDM II/Labs/Lab 06_Buffers")
#dir.create( "shapefiles" )
setwd( "./shapefiles" )


# Download shapefiles
download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2010/TRACT/2010/tl_2010_36067_tract10.zip", "onondaga census tracts.zip" )
unzip( "onondaga census tracts.zip" )
file.remove( "onondaga census tracts.zip" )
dir()


# Load and plot shapefiles
syr <- readShapePoly( fn="tl_2010_36067_tract10", proj4string=CRS("+proj=longlat +datum=WGS84") )

par( mar=c(0,0,0,0) )
plot( syr,  border="gray10" )


# Add roads

download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2015/PRISECROADS/tl_2015_36_prisecroads.zip", "primary_and_secondary_roads.zip" )

unzip( "primary_and_secondary_roads.zip" )
file.remove( "primary_and_secondary_roads.zip" )
dir()

roads <- readShapeLines( fn="tl_2015_36_prisecroads", proj4string=CRS("+proj=longlat +datum=WGS84") )

head( roads@data )
table( roads$RTTYP )
class( roads )
#RTTYP = Road Type Code
# C = county
# I = interstate
# M = Common Name
# O = other
# S = state recognized
# U = U.S.
```

#### Create a polygon outline
```{r}
head( syr@data )
names( syr )
table( syr$GEOID10 )
str( syr$TRACTCE10 )

# Plot only the tracts within the city
# 0 or 1 in the 7th position of the FIPS ID string seem to differentiate city vs rural
# OR only looking at the last 4 is insufficient for a 3-digit census tract ID

last5tract <- unique(substr(syr$GEOID10, 7,11))
table(last5tract)
syr.sub <- syr[ (as.numeric(substr(syr$GEOID10, 7,11)) < 06200 ) & (as.numeric(substr(syr$GEOID10, 7,11)) > 0), ]
head(syr.sub)
plot(syr.sub)
class( syr.sub )
```


#### Clip Roads down to same size as census tracts
```{r}
# Using gClip() to do a modified subsetting of geospatial points/lines
# http://robinlovelace.net/r/2014/07/29/clipping-with-r.html
# rgeos version: 0.3-5, (SVN revision 447)
#  GEOS runtime version: 3.4.2-CAPI-1.8.2 r3921 
#  Polygon checking: TRUE

gClip <- function(shp, bb){  #shp = shapefile, bb = bounding box
  if(class(bb) == "matrix") b_poly <- as(extent(as.vector(t(bb))), "SpatialPolygons")
  else b_poly <- as(extent(bb), "SpatialPolygons")
  gIntersection(shp, b_poly, byid = T)
}

roads.clipped <- gClip(roads, syr)
#this returns a BOX that has the roads limited within it
plot( roads.clipped )
str( roads.clipped )
class( roads.clipped )
# Clip is computationally expensive, has to determine what to do with the polygons destroyed in the clip

# use intersect instead of gClip and syr.sub
## source: https://gis.stackexchange.com/questions/23931/how-to-intersect-lines-and-polygons-in-r
## author: xhie                library(raster)
## newLines <- raster::intersect(SpatialLinesDataframe, SpatialPolygonDataframe)
roads.cropped <- raster::intersect( roads, syr.sub )
plot( roads.cropped )
str( roads.cropped )
class( roads.cropped )
table( roads.cropped$RTTYP )

```


#### Extract the interstates as a separate layer
```{r}
#subset()

# Using roads.clipped

par( mar=c(0,0,0,0) )
plot( syr,  border="gray80" )
#plot( roads, col="steelblue", lwd=2, add=T )
plot( roads.clipped, col="steelblue", lwd=1, add=T )

#interstate <- roads[ roads$RTTYP == "I", ]
interstate1 <- roads.clipped[ roads.clipped$RTTYP == "I", ]
interstate.clipped <- gClip( interstate, syr )
#this returns a BOX that has the interstates limited within it
plot( interstate.clipped, col="red", add=T )



# Using roads.cropped

par( mar=c(0,0,0,0) )
plot( syr.sub,  border="gray80" )
#plot( roads, col="steelblue", lwd=2, add=T )
plot( roads.cropped, col="steelblue", lwd=1, add=T )

#interstate <- roads[ roads$RTTYP == "I", ]
interstate2 <- roads.cropped[ roads.cropped$RTTYP == "I", ]
interstate.cropped <- raster::intersect( interstate2, syr.sub )

plot( interstate.cropped, col="red", add=T )
class( interstate.cropped )
```

#### Create a buffer of approximately a quarter mile from the interstate
```{r}
#interstate.projected <- spTransform( interstate.cropped, CRS( "+init=epsg:3347" ) ) 
#coordinates(interstate.projected)

#class( interstate.projected )

# Approximate Buffer
buffhwy <- gBuffer(
  interstate.cropped,
  #interstate.projected,
  #width = 0.001, # width of buffer
  width = 0.005, # width of buffer
  byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
)
class( buffhwy )
# buffhwy <- SpatialPolygonsDataFrame( buffhwy, data=buffhwy@data)
class( buffhwy )
str( buffhwy )
plot( buffhwy, add=T)


# Accurate Buffer
# To make accurate buffers, use a variant of this code.
# Units: https://gis.stackexchange.com/questions/85300/units-of-gbuffer-from-rgeos-package
# Projection to use: https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/OverviewCoordinateReferenceSystems.pdf
interstate.projected.US <- spTransform( 
  interstate.cropped, # source polygon
  CRS( 
    "+proj=utm +zone=18 +datum=WGS84" ) # use the UTM projection, Syracuse is in Zone 18
  )

unit <- conv_unit( 10, "mi", "m" ) # converts from 0.25 mi to meters, the unit of UTM
buffhwy2 <- gBuffer(
    interstate.projected.US,
    #interstate.projected,
    #width = 0.001, # width of buffer
    width = unit,
    byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
  )
class( buffhwy2 )
# buffhwy <- SpatialPolygonsDataFrame( buffhwy, data=buffhwy@data)
str( buffhwy2 )
plot( buffhwy2, add=T)

```

#### Identify all houses that fall within the buffer zone
```{r}
zdat <- read.csv( "https://raw.githubusercontent.com/R-Class/all-labs-ddmii-fjsantam/master/Lab%2006/zillowdata.csv" )
# For the code used to generate the raw CSV file above, refer to Lab_06_-_Geocoding_CSV_Script_FJS.Rmd in the Lab 06 folder on GitHub
names( zdat )
nrow( zdat )

# Convert the coordinates that are in lon (x) lat(y) order to spatial points, identifying the coordinates, data, and coordinate system
zcoor <- zdat[ c( "lon", "lat") ]
spdat <- SpatialPointsDataFrame( zcoor, zdat, proj4string=CRS("+proj=longlat +datum=WGS84")  )
class( spdat )
head( spdat )
#points( spdat )
nrow( spdat )

# Function is: over( locations, layer )
homes.on.highway <- over( spdat, buffhwy )
class( homes.on.highway)

# buff1 is a SpatialPolygon; dat is a dataframe
# will return a list of points (top row) and which polygon they are in (represented by ID # or NA if not in one)
honh.logical <- homes.on.highway[is.na(homes.on.highway)]

honh.logical <- as.character(homes.on.highway) != "NA"
honh.logical[ is.na(honh.logical) ] <- FALSE
class( honh.logical )
```


#### Add a new categorical variable to the houses dataset: within buffer zone or not?
```{r}
dat <- as.data.frame( zdat )
dat.buffer <- cbind( dat, honh.logical )
head( dat.buffer )
nrow( dat.buffer )

dat.hwy.buffer <- dat.buffer[ dat.buffer$honh.logical == TRUE, ]
nrow( dat.hwy.buffer )
head(dat.hwy.buffer)
```


#### Compile the Map
```{r}
par( mar=c(0,0,0,0) )
plot( syr.sub,  border="gray80" )
plot( roads.cropped, col="steelblue", lwd=1, add=T )
#plot( buffhwy, add=T)
plot( interstate.cropped, col="red", add=T )
points( spdat, cex = 0.5 )
points( x=dat.hwy.buffer$lon, y=dat.hwy.buffer$lat, pch=19, col="magenta" )
```


## Part 2

#### Create a buffer a quarter mile from industrial zones (LandUse) and plot the buffer zone
#### Redo this using the original map file and geocoding addresses in the parcels file, instead of the syr2 shapefile
*Note*: To get an accurate buffer zone away from industrial zones, the polygon would have to be read in as a projection of UTM. But the data is in long-lat coordinates. I am unsure how to convert long-lat coordinates for a property to UTM coordinates. Without this conversion, the buffer will have to be an estimate rather than an accurate representation.

Issues:
```{r}
#parcels <- read.csv( "https://raw.githubusercontent.com/lecy/maps-in-R/master/Data/syr_parcels.csv" )
#names(parcels)
#unique( parcels$LandUse )

### Get census tracts FIPS, grab all ids of 1-61, dissolve the polygons to get the outline
setwd( "~/Graduate School/PAI 690 Independent Study_DDM II/Labs/shapefiles" )
# download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2010/TRACT/2010/tl_2010_36067_tract10.zip", "onondaga census tracts.zip" )
# unzip( "onondaga census tracts.zip" )
# file.remove( "onondaga census tracts.zip" )
dir()
syr2 <- readShapePoly( fn="01-05-2015", 
         #Approximate:
         proj4string=CRS("+proj=longlat +datum=WGS84") 
         #Accurate:
         #proj4string=CRS( "+proj=utm +zone=18 +datum=WGS84" ) 
         )
head( syr2@data) #returns the output of a list of values in syr2 map data
class(syr2)

landUseInd <- syr2[ syr2$LandUse == "Industrial", ]
class(landUseInd)
landUseNotInd <- syr2[ syr2$LandUse != "Industrial", ]
#unique(syr2$LandUse)

par( mar=c(0,0,0,0) )
plot( landUseNotInd, col = "grey", border = F )
plot( landUseInd, col = "orange", border = F, add = T )
```

Converting 
```{r}
# https://stackoverflow.com/questions/18639967/converting-latitude-and-longitude-points-to-utm
# See the answer posted by Stanislav

LongLatToUTM <- function( x, y, zone){
  xy <- data.frame( ID = 1:length(x), X = x, Y = y)
  coordinates( xy ) <- c( "X", "Y" )
  proj4string( xy ) <- CRS( "+proj=longlat +datum=WGS84" )
  res <- spTransform( xy, CRS( paste( "+proj=utm +zone=",zone,"ellps=WGS84", sep = '') ) )
  return( as.data.frame( res ) )
}
# Create a vector of the UTM values
spUTMdat <- LongLatToUTM( spdat$lon, spdat$lat, 18 )
class(spUTMdat)
head( spUTMdat )
nrow(spUTMdat)

# Transform the vector into an object of class SpatialPoints
spUTMcoor <- SpatialPoints(
  spUTMdat, 
  CRS( "+proj=utm +zone=18 +datum=WGS84" ) # use the UTM projection, Syracuse is in Zone 18
  )
class( spUTMcoor)


syr2acc <- readShapePoly( fn="01-05-2015", 
         #Approximate:
         #proj4string=CRS("+proj=longlat +datum=WGS84") 
         #Accurate:
         proj4string=CRS( "+proj=utm +zone=18 +datum=WGS84" ) 
         )
head( syr2acc@data) #returns the output of a list of values in syr2 map data
class(syr2acc)

landUseIndacc <- syr2acc[ syr2acc$LandUse == "Industrial", ]
class(landUseIndacc)
```

##### Creating Buffers
```{r}
# Approximate Buffer

syr2 <- spTransform( syr2, CRS)

buffindzone <- gBuffer(
  landUseInd,
  # width = 0.001, # width of buffer
  # width = 400, # width of buffer
  # width = 2400, # expanding to play nice with the 189-house dataset
  #width = 0.003621,
  width = 3600,
  capStyle = "FLAT",
  quadsegs = 1,
  byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
)
class( buffindzone )

# Accurate Buffer

buffindzoneacc <- gBuffer(
  landUseIndacc,
  #width = 0.001, # width of buffer
  #width = 400, # width of buffer
  width = 2400, # expanding to play nice with the 189-house dataset
  byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
)
class( buffindzoneacc )


#plot( syr2,  border="gray80" )
plot( buffindzone, add=T, col=rgb(44,4,248,alpha=50, maxColorValue = 355))
map.scale( metric=F, ratio=F, relwidth=0.15, cex=0.5)

```

# Create a vector that contains the addresses present in both the Syracuse polygon and the Zillow data
```{r}
# https://stackoverflow.com/questions/16516593/convert-from-lowercase-to-uppercase-all-values-in-all-character-variables-in-dat
# See answer posted by juba

upCase <- function(v) {
  if( is.character(v)) return(toupper(v))
  else return(v)
}


test.syr2 <- c( syr2$Add3, syr2$Add4, syr2$ZIP)
#test.syr2 <- c( syr2$Add3, syr2$Add4 )
test.syr2 <- paste( syr2$Add3, syr2$Add4, syr2$ZIP, sep=", " )
#test.syr2 <- paste( syr2$Add3, syr2$Add4, sep=", " )
class( test.syr2 )
head( test.syr2 )
test.spdat <- paste( spdat$address, "Syracuse  NY", spdat$zip, sep=", " )
#test.spdat <- paste( spdat$address, "Syracuse  NY", sep=", " )
test.spdat <- upCase( test.spdat )
head( test.spdat )

#data.frame( lapply( df, upCase(test.spdat) ) )

test.match <- match( upCase( as.character( spdat$address ) ), as.character( syr2$Add3 ) )
test.match <- match( test.spdat, test.syr2 )
table( test.match )
summary(test.match)



test2.syr2 <- syr2$Add3
test2.syr2 <- as.character(test2.syr2)
head(test2.syr2)

test2.spdat <- spdat$address
test2.spdat <- as.character( test2.spdat)
test2.spdat <- upCase( test2.spdat )
head(test2.spdat)


test2.match <- match( test2.spdat, test2.syr2 )
table(test2.match)
test2.merge <- merge( spdat, syr2, by.x = upCase(as.character( spdat$address ) ), by.y = as.character( syr2$Add3 )  )
head(test2.merge)




## Create a subset of syr2 for only the addresses that are found in spdat
test2.spdat <- as.factor(test2.spdat)
class(test2.spdat)
head(test2.spdat)

spdat.syr2 <- syr2[ syr2$Add3 %in% test2.spdat, ]
cleanedAddresses <- spdat.syr2$Add3
# spdat.syr2 <- match( upCase(as.character( spdat$address ) ), as.character( syr2$Add3 ) )
head( spdat.syr2 )
nrow( spdat.syr2 )

```



#### Identify houses within the buffer zones and create a categorical variable in the dataset indicating proximity to industrial zones
```{r}
  
points( x = spdat$lon, y = spdat$lat, pch = 19 )


#Accurate Version

# Convert the housing data to use the same CRS system
spdat2 <- spTransform( 
  spdat, # source polygon
  CRS( "+proj=utm +zone=18 +datum=WGS84" ) # use the UTM projection, Syracuse is in Zone 18
  )
head(spdat2)
nrow(spdat2)

# Function is: over( locations, layer )
#homes.in.ind <- over( spdat2, buffindzone )
homes.in.ind <- over( spUTMcoor, buffindzone )
class( homes.in.ind)

# buff1 is a SpatialPolygon; dat is a dataframe
# will return a list of points (top row) and which polygon they are in (represented by ID # or NA if not in one)
hini.logical <- homes.in.ind[is.na(homes.in.ind)]

hini.logical <- as.character(homes.in.ind) != "NA"
hini.logical[ is.na(hini.logical) ] <- FALSE
class( hini.logical )

# Approximate Version

# Function is: over( locations, layer )
homes.in.ind <- over( spdat.syr2, buffindzone )
class( homes.in.ind)

# buff1 is a SpatialPolygon; dat is a dataframe
# will return a list of points (top row) and which polygon they are in (represented by ID # or NA if not in one)
hini.logical <- homes.in.ind[is.na(homes.in.ind)]

hini.logical <- as.character(homes.in.ind) != "NA"
hini.logical[ is.na(hini.logical) ] <- FALSE
class( hini.logical )



############# Add to the dataset #############
proper=function(s) gsub("(?<=\\b)([a-z])","\\U\\1",tolower(s), perl=T)
cleanedAddresses <- proper(cleanedAddresses)
cleanedAddresses <- as.factor(cleanedAddresses)
dat.buffer.syr2 <- dat.buffer[ upCase(dat.buffer$address) %in% cleanedAddresses, ]
dat.buffer <- cbind( dat.buffer.syr2, hini.logical )
head( dat.buffer )
nrow( dat.buffer )

dat.ind.buffer <- dat.buffer[ dat.buffer$hini.logical == TRUE, ]
nrow( dat.ind.buffer )
head(dat.ind.buffer)
dat.not.ind.buffer <- dat.buffer[ dat.buffer$hini.logical == FALSE, ]
nrow( dat.not.ind.buffer )


# Plot the map, overlaying roads, houses, and houses within the industrial zones
plot( syr2,  border="gray80" )
plot( roads.cropped, col="steelblue", lwd=1, add=T )
#plot( buffhwy, add=T)
points( x=spdat2$lon, y=spdat2$lat, cex = 0.5, col = "grey" )
points( x=dat.ind.buffer$lon, y=dat.ind.buffer$lat, pch=19, col="orange" )
# points( x=dat.not.ind.buffer$lon, y=dat.not.ind.buffer$lat, pch=19, col="grey" )

```

#### Create a buffer zone an eighth of a mile from schools, and plot the buffer zone.
```{r}

```

#### Identify houses within the buffer zones and create a categorical variable in the dataset indicating proximity to schools
```{r}

```

