---
title: "Lab 06 - Creating Buffers"
author: "Francisco Santamarina"
date: "March 09, 2017"
output:
  html_document:
    df_print: paged
    keep_md: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F )
```

Load the necessary packages.
```{r}
#library( censusAPI)
library( geojsonio )
#library( ggmap )
#library( jsonlite )
library( maps )
library( maptools )
library( measurements )
#library( memisc )
#library( RCurl )
library( raster )
library( rgdal )
library( rgeos )
library( sp )
#library( spatialEco ) 

```

## Part 1: Tracts with TIGER Files

Load and clean the necessary dataset.
```{r}
# Set the working directory
#setwd( "//hd.ad.syr.edu/02/7796a1/Documents/Desktop/DDM/DDM II/Lab 06" )
setwd("~/Graduate School/PAI 690 Independent Study_DDM II/Labs/Lab 06_Buffers")
#dir.create( "shapefiles" )
setwd( "./shapefiles" )


# Download shapefiles
download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2010/TRACT/2010/tl_2010_36067_tract10.zip", "onondaga census tracts.zip" )
unzip( "onondaga census tracts.zip" )
file.remove( "onondaga census tracts.zip" )
dir()


# Load and plot shapefiles
syr <- readShapePoly( fn="tl_2010_36067_tract10", proj4string=CRS("+proj=longlat +datum=WGS84") )

par( mar=c(0,0,0,0) )
plot( syr,  border="gray10" )


# Add roads

download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2015/PRISECROADS/tl_2015_36_prisecroads.zip", "primary_and_secondary_roads.zip" )

unzip( "primary_and_secondary_roads.zip" )
file.remove( "primary_and_secondary_roads.zip" )
dir()

roads <- readShapeLines( fn="tl_2015_36_prisecroads", proj4string=CRS("+proj=longlat +datum=WGS84") )

head( roads@data )
table( roads$RTTYP )
class( roads )
#RTTYP = Road Type Code
# C = county
# I = interstate
# M = Common Name
# O = other
# S = state recognized
# U = U.S.
```

#### Create a polygon outline
```{r}
head( syr@data )
names( syr )
table( syr$GEOID10 )
str( syr$TRACTCE10 )

# Plot only the tracts within the city
# 0 or 1 in the 7th position of the FIPS ID string seem to differentiate city vs rural
# OR only looking at the last 4 is insufficient for a 3-digit census tract ID

last5tract <- unique(substr(syr$GEOID10, 7,11))
table(last5tract)
syr.sub <- syr[ (as.numeric(substr(syr$GEOID10, 7,11)) < 06200 ) & (as.numeric(substr(syr$GEOID10, 7,11)) > 0), ]
head(syr.sub)
plot(syr.sub)
class( syr.sub )
```


#### Clip Roads down to same size as census tracts
```{r}
# Using gClip() to do a modified subsetting of geospatial points/lines
# http://robinlovelace.net/r/2014/07/29/clipping-with-r.html
# rgeos version: 0.3-5, (SVN revision 447)
#  GEOS runtime version: 3.4.2-CAPI-1.8.2 r3921 
#  Polygon checking: TRUE

gClip <- function(shp, bb){  #shp = shapefile, bb = bounding box
  if(class(bb) == "matrix") b_poly <- as(extent(as.vector(t(bb))), "SpatialPolygons")
  else b_poly <- as(extent(bb), "SpatialPolygons")
  gIntersection(shp, b_poly, byid = T)
}

roads.clipped <- gClip(roads, syr)
#this returns a BOX that has the roads limited within it
plot( roads.clipped )
str( roads.clipped )
class( roads.clipped )
# Clip is computationally expensive, has to determine what to do with the polygons destroyed in the clip

# use intersect instead of gClip and syr.sub
## source: https://gis.stackexchange.com/questions/23931/how-to-intersect-lines-and-polygons-in-r
## author: xhie                library(raster)
## newLines <- raster::intersect(SpatialLinesDataframe, SpatialPolygonDataframe)
# raster::intersect is a way to access intersect within package raster. Loads the package without attaching it.
roads.cropped <- intersect( roads, syr.sub )
plot( roads.cropped )
str( roads.cropped )
class( roads.cropped )
table( roads.cropped$RTTYP )

```


#### Extract the interstates as a separate layer
```{r}
#subset()

# Using roads.clipped

par( mar=c(0,0,0,0) )
plot( syr,  border="gray80" )
#plot( roads, col="steelblue", lwd=2, add=T )
plot( roads.clipped, col="steelblue", lwd=1, add=T )

#interstate <- roads[ roads$RTTYP == "I", ]
interstate1 <- roads.clipped[ roads.clipped$RTTYP == "I", ]
interstate.clipped <- gClip( interstate, syr )
#this returns a BOX that has the interstates limited within it
plot( interstate.clipped, col="red", add=T )



# Using roads.cropped

par( mar=c(0,0,0,0) )
plot( syr.sub,  border="gray80" )
#plot( roads, col="steelblue", lwd=2, add=T )
plot( roads.cropped, col="steelblue", lwd=1, add=T )

#interstate <- roads[ roads$RTTYP == "I", ]
interstate2 <- roads.cropped[ roads.cropped$RTTYP == "I", ]
interstate.cropped <- intersect( interstate2, syr.sub )

plot( interstate.cropped, col="red", add=T )
class( interstate.cropped )
```

#### Create a buffer of approximately a quarter mile from the interstate
```{r}
#interstate.projected <- spTransform( interstate.cropped, CRS( "+init=epsg:3347" ) ) 
#coordinates(interstate.projected)

#class( interstate.projected )

# Approximate Buffer
buffhwy <- gBuffer(
  interstate.cropped,
  #interstate.projected,
  #width = 0.001, # width of buffer
  width = 0.005, # width of buffer
  byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
)
class( buffhwy )
# buffhwy <- SpatialPolygonsDataFrame( buffhwy, data=buffhwy@data)
str( buffhwy )
plot( buffhwy, add=T)


# Accurate Buffer
# To make accurate buffers, use a variant of this code.
# Units: https://gis.stackexchange.com/questions/85300/units-of-gbuffer-from-rgeos-package
# Projection to use: https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/OverviewCoordinateReferenceSystems.pdf
interstate.projected.US <- spTransform( 
  interstate.cropped, # source polygon
  CRS( 
    "+proj=utm +zone=18 +datum=WGS84" ) # use the UTM projection, Syracuse is in Zone 18
  )

unit <- conv_unit( 10, "mi", "m" ) # converts from 0.25 mi to meters, the unit of UTM
buffhwy2 <- gBuffer(
    interstate.projected.US,
    #interstate.projected,
    #width = 0.001, # width of buffer
    width = unit,
    byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
  )
class( buffhwy2 )
# buffhwy <- SpatialPolygonsDataFrame( buffhwy, data=buffhwy@data)
str( buffhwy2 )
plot( buffhwy2, add=T)

```

#### Identify all houses that fall within the buffer zone
```{r}
zdat <- read.csv( "https://raw.githubusercontent.com/R-Class/all-labs-ddmii-fjsantam/master/Lab%2006/zillowdata.csv" )
# For the code used to generate the raw CSV file above, refer to Lab_06_-_Geocoding_CSV_Script_FJS.Rmd in the Lab 06 folder on GitHub
names( zdat )
nrow( zdat )

# Convert the coordinates that are in lon (x) lat(y) order to spatial points, identifying the coordinates, data, and coordinate system
zcoor <- zdat[ c( "lon", "lat") ]
spdat <- SpatialPointsDataFrame( zcoor, zdat, proj4string=CRS("+proj=longlat +datum=WGS84")  )
class( spdat )
head( spdat )
#points( spdat )
nrow( spdat )

# Function is: over( locations, layer )
homes.on.highway <- over( spdat, buffhwy )
class( homes.on.highway)

# buff1 is a SpatialPolygon; dat is a dataframe
# will return a list of points (top row) and which polygon they are in (represented by ID # or NA if not in one)
honh.logical <- homes.on.highway[is.na(homes.on.highway)]

honh.logical <- as.character(homes.on.highway) != "NA"
honh.logical[ is.na(honh.logical) ] <- FALSE
class( honh.logical )
```


#### Add a new categorical variable to the houses dataset: within buffer zone or not?
```{r}
dat <- as.data.frame( zdat )
dat.buffer <- cbind( dat, honh.logical )
head( dat.buffer )
nrow( dat.buffer )

dat.hwy.buffer <- dat.buffer[ dat.buffer$honh.logical == TRUE, ]
nrow( dat.hwy.buffer )
head(dat.hwy.buffer)
```


#### Compile the Map
```{r}
par( mar=c(0,0,0,0) )
plot( syr.sub,  border="gray80" )
plot( roads.cropped, col="steelblue", lwd=1, add=T )
#plot( buffhwy, add=T)
plot( interstate.cropped, col="red", add=T )
points( spdat, cex = 0.5 )
points( x=dat.hwy.buffer$lon, y=dat.hwy.buffer$lat, pch=19, col="magenta" )
```



## Part 2: Parcels with GeoJSON Files

#### Create a buffer a quarter mile from industrial zones (LandUse) and plot the buffer zone

```{r}
# Set the working directory
#setwd( "//hd.ad.syr.edu/02/7796a1/Documents/Desktop/DDM/DDM II/Lab 06" )
setwd("~/GitHub/all-labs-ddmii-fjsantam/Lab 06")
#dir.create( "shapefiles" )
setwd( "./shapefiles" )


# Download shapefiles
# download.file("https://github.com/lecy/geojson/raw/master/syr_parcels.geojson", "syr_parcels.geojson" )
dir()

######## Shout out to Christopher Davis and his code
# Load and plot shapefiles
syr2 <- readOGR( dsn = "shapefiles/syr_parcels.geojson" )

plot( syr2 )
head( syr2@data)
names( syr2 )
#unique( parcels$LandUse )

landUseInd <- syr2[ syr2$LandUse == "Industrial", ]
landUseNotInd <- syr2[ syr2$LandUse != "Industrial", ]
#unique(syr2$LandUse)

par( mar=c(0,0,0,0) )
plot( landUseNotInd, col = "grey", border = F )
plot( landUseInd, col = "orange", border = F, add = T )

```

##### Creating Buffers
```{r}
# Approximate Buffer

buffindzone <- gBuffer(
  landUseInd,
  # width = 0.001, # width of buffer
  # width = 400, # width of buffer
  # width = 2400, # expanding to play nice with the 189-house dataset
  width = 0.003621,
  #width = 3600,
  capStyle = "FLAT",
  quadsegs = 1,
  byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
)


#plot( syr2,  border="gray80" )
plot( buffindzone, add=T, col=rgb(44,4,248,alpha=50, maxColorValue = 355))
map.scale( metric=F, ratio=F, relwidth=0.15, cex=0.5)

```


##### Identify houses within the buffer zones and create a categorical variable in the dataset indicating proximity to industrial zones
```{r}
zdat <- read.csv( "https://raw.githubusercontent.com/R-Class/all-labs-ddmii-fjsantam/master/Lab%2006/zillowdata.csv", stringsAsFactors = F )
zcoor <- zdat[ c( "lon", "lat") ]
spdat <- SpatialPoints( zcoor, proj4string=CRS("+proj=longlat +datum=WGS84")  )


#spdat <- spTransform( spdat, CRS("+proj=longlat +datum=WGS84 +towgs84=0,0,0 +ellps=WGS84 +no_defs") )
buffindzone <- spTransform( buffindzone, CRS( "+proj=longlat +datum=WGS84" ) )

# Function is: over( locations, layer )
homes.in.ind <- over( spdat, buffindzone )
class( homes.in.ind)

# buff1 is a SpatialPolygon; dat is a dataframe
# will return a list of points (top row) and which polygon they are in (represented by ID # or NA if not in one)
#hini.logical <- homes.in.ind[is.na(homes.in.ind)]

hini.logical <- as.character(homes.in.ind) != "NA"
hini.logical[ is.na(hini.logical) ] <- FALSE
class( hini.logical )



############# Add to the dataset #############

dat.buffer <- cbind( dat.buffer, hini.logical )
head( dat.buffer )
nrow( dat.buffer )

dat.ind.buffer <- dat.buffer[ dat.buffer$hini.logical == TRUE, ]
nrow( dat.ind.buffer )
head(dat.ind.buffer)
#dat.not.ind.buffer <- dat.buffer[ dat.buffer$hini.logical == FALSE, ]
#nrow( dat.not.ind.buffer )


# Plot the map, overlaying roads, houses, and houses within the industrial zones
#plot( syr2,  border="gray80" )
par( mar=c(0,0,0,0) )
plot( landUseNotInd, col = "grey", border = F )
plot( landUseInd, col = "orange", border = F, add = T )
plot( buffindzone, add=T )
#plot( roads.cropped, col="steelblue", lwd=1, add=T )
#plot( buffindzone, add=T)
points( x=spdat$lon, y=spdat$lat, cex = 0.5, pch=1 #col = "grey" 
       )
points( x=dat.ind.buffer$lon, y=dat.ind.buffer$lat, pch=19, col="goldenrod4" )
# points( x=dat.not.ind.buffer$lon, y=dat.not.ind.buffer$lat, pch=19, col="grey" )
map.scale( metric=F, ratio=F, relwidth=0.15, cex=0.5)
```

#### Create a buffer zone an eighth of a mile from schools, and plot the buffer zone.
```{r}
# Identify the schools
#unique(syr2$LandUse)
landUseSchool <- syr2[ syr2$LandUse == "Schools", ]
landUseNotSchool <- syr2[ syr2$LandUse != "Schools", ]

# Create the buffer
buffschzone <- gBuffer(
  landUseSchool,
  # width = 0.001, # width of buffer
  # width = 400, # width of buffer
  # width = 2400, # expanding to play nice with the 189-house dataset
  #width = 0.003621,
  width = 0.0018105,
  #width = 3600,
  capStyle = "FLAT",
  quadsegs = 1,
  byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
)

par( mar=c(0,0,0,0) )
plot( landUseNotSchool, col = "grey", border = F )
plot( landUseSchool, col = "violetred1", border = F, add = T )
plot( buffschzone, add=T )
map.scale( metric=F, ratio=F, relwidth=0.15, cex=0.5)

```

#### Identify houses within the buffer zones and create a categorical variable in the dataset indicating proximity to schools
```{r}
buffschzone <- spTransform( buffschzone, CRS( "+proj=longlat +datum=WGS84" ) )
homes.in.sch <- over( spdat, buffschzone )
class( homes.in.sch)

# buff1 is a SpatialPolygon; dat is a dataframe
# will return a list of points (top row) and which polygon they are in (represented by ID # or NA if not in one)
#hins.logical <- homes.in.sch[is.na(homes.in.sch)]

hins.logical <- as.character(homes.in.sch) != "NA"
hins.logical[ is.na(hins.logical) ] <- FALSE
class( hins.logical )



############# Add to the dataset #############

dat.buffer <- cbind( dat.buffer, hins.logical )
head( dat.buffer )
nrow( dat.buffer )

dat.sch.buffer <- dat.buffer[ dat.buffer$hini.logical == TRUE, ]
nrow( dat.sch.buffer )
head(dat.sch.buffer)
#dat.not.ind.buffer <- dat.buffer[ dat.buffer$hini.logical == FALSE, ]
#nrow( dat.not.ind.buffer )


# Plot the map, overlaying roads, houses, and houses within the industrial zones
#plot( syr2,  border="gray80" )
par( mar=c(0,0,0,0) )
plot( landUseNotSchool, col = "grey", border = F )
plot( landUseSchool, col = "violetred1", border = F, add = T )
#plot( roads.cropped, col="steelblue", lwd=1, add=T )
#plot( buffindzone, add=T)
points( x=spdat$lon, y=spdat$lat, cex = 0.5, pch=1 #col = "grey" 
       )
points( x=dat.sch.buffer$lon, y=dat.sch.buffer$lat, pch=19, col="springgreen2" )
# points( x=dat.not.ind.buffer$lon, y=dat.not.ind.buffer$lat, pch=19, col="grey" )
plot( buffschzone, add=T )
map.scale( metric=F, ratio=F, relwidth=0.15, cex=0.5)

```

