---
title: "Lab 06 - Creating Buffers"
author: "Francisco Santamarina"
date: "March 09, 2017"
output:
  html_document:
    df_print: paged
    keep_md: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F )
```

## Part 1: Tracts

Load the necessary packages.
```{r}
#library( censusAPI)
#library( ggmap )
#library( jsonlite )
library( maps )
library( maptools )
library( measurements )
#library( memisc )
#library( RCurl )
library( raster )
library( rgdal )
library( rgeos )
library( sp )
#library( spatialEco ) 

```

Load and clean the necessary dataset.
```{r}
# Set the working directory
#setwd( "//hd.ad.syr.edu/02/7796a1/Documents/Desktop/DDM/DDM II/Lab 06" )
setwd("~/Graduate School/PAI 690 Independent Study_DDM II/Labs/Lab 06_Buffers")
#dir.create( "shapefiles" )
setwd( "./shapefiles" )


# Load and plot shapefiles
syr <- readShapePoly( fn="tl_2010_36067_tract10", proj4string=CRS("+proj=longlat +datum=WGS84") )

# par( mar=c(0,0,0,0) )
# plot( syr,  border="gray10" )


# Add roads

roads <- readShapeLines( fn="tl_2015_36_prisecroads", proj4string=CRS("+proj=longlat +datum=WGS84") )
```

#### Create a polygon outline
```{r}
syr.sub <- syr[ (as.numeric(substr(syr$GEOID10, 7,11)) < 06200 ) & (as.numeric(substr(syr$GEOID10, 7,11)) > 0), ]
```


#### Clip Roads down to same size as census tracts
```{r}
roads.cropped <- raster::intersect( roads, syr.sub )

```


#### Extract the interstates as a separate layer
```{r}
#interstate <- roads[ roads$RTTYP == "I", ]
interstate2 <- roads.cropped[ roads.cropped$RTTYP == "I", ]
interstate.cropped <- raster::intersect( interstate2, syr.sub )

```

#### Create a buffer of approximately a quarter mile from the interstate
```{r}

# Approximate Buffer
buffhwy <- gBuffer(
  interstate.cropped,
  #interstate.projected,
  #width = 0.001, # width of buffer
  width = 0.005, # width of buffer
  byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
)

```

#### Identify all houses that fall within the buffer zone
```{r}
zdat <- read.csv( "https://raw.githubusercontent.com/R-Class/all-labs-ddmii-fjsantam/master/Lab%2006/zillowdata.csv" )
# For the code used to generate the raw CSV file above, refer to Lab_06_-_Geocoding_CSV_Script_FJS.Rmd in the Lab 06 folder on GitHub

# Convert the coordinates that are in lon (x) lat(y) order to spatial points, identifying the coordinates, data, and coordinate system
zcoor <- zdat[ c( "lon", "lat") ]
spdat <- SpatialPointsDataFrame( zcoor, zdat, proj4string=CRS("+proj=longlat +datum=WGS84")  )

# Function is: over( locations, layer )
homes.on.highway <- over( spdat, buffhwy )

# buff1 is a SpatialPolygon; dat is a dataframe
# will return a list of points (top row) and which polygon they are in (represented by ID # or NA if not in one)
honh.logical <- homes.on.highway[is.na(homes.on.highway)]

honh.logical <- as.character(homes.on.highway) != "NA"
honh.logical[ is.na(honh.logical) ] <- FALSE
```


#### Add a new categorical variable to the houses dataset: within buffer zone or not?
```{r}
dat <- as.data.frame( zdat )
dat.buffer <- cbind( dat, honh.logical )

dat.hwy.buffer <- dat.buffer[ dat.buffer$honh.logical == TRUE, ]
```


#### Compile the Map
```{r}
par( mar=c(0,0,0,0) )
plot( syr.sub,  border="gray80" )
plot( roads.cropped, col="steelblue", lwd=1, add=T )
#plot( buffhwy, add=T)
plot( interstate.cropped, col="red", add=T )
points( spdat, cex = 0.5 )
points( x=dat.hwy.buffer$lon, y=dat.hwy.buffer$lat, pch=19, col="magenta" )
```


## Part 2: Parcels

#### Create a buffer a quarter mile from industrial zones (LandUse) and plot the buffer zone
#### Redo this using the original map file and geocoding addresses in the parcels file, instead of the syr2 shapefile
*Note*: To get an accurate buffer zone away from industrial zones, the polygon would have to be read in as a projection of UTM. But the data is in long-lat coordinates. I am unsure how to convert long-lat coordinates for a property to UTM coordinates. Without this conversion, the buffer will have to be an estimate rather than an accurate representation.

Issues:
```{r}
#parcels <- read.csv( "https://raw.githubusercontent.com/lecy/maps-in-R/master/Data/syr_parcels.csv" )
#names(parcels)
#unique( parcels$LandUse )

### Get census tracts FIPS, grab all ids of 1-61, dissolve the polygons to get the outline
setwd( "~/Graduate School/PAI 690 Independent Study_DDM II/Labs/shapefiles" )
# download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2010/TRACT/2010/tl_2010_36067_tract10.zip", "onondaga census tracts.zip" )
# unzip( "onondaga census tracts.zip" )
# file.remove( "onondaga census tracts.zip" )
syr2 <- readShapePoly( fn="01-05-2015", 
         #Approximate:
         proj4string=CRS("+proj=longlat +datum=WGS84") 
         #Accurate:
         #proj4string=CRS( "+proj=utm +zone=18 +datum=WGS84" ) 
         )

landUseInd <- syr2[ syr2$LandUse == "Industrial", ]
landUseNotInd <- syr2[ syr2$LandUse != "Industrial", ]
#unique(syr2$LandUse)

```

##### Creating Buffers
```{r}
# Approximate Buffer

buffindzone <- gBuffer(
  landUseInd,
  # width = 0.001, # width of buffer
  # width = 400, # width of buffer
  # width = 2400, # expanding to play nice with the 189-house dataset
  #width = 0.003621,
  width = 2400,
  capStyle = "FLAT",
  quadsegs = 1,
  byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
)
class( buffindzone )

buffindzone <- spTransform( 
  buffindzone, # source polygon
  CRS( "+proj=longlat +datum=WGS84" )
  )


#plot( syr2,  border="gray80" )

#map.scale( metric=F, ratio=F, relwidth=0.15, cex=0.5)

```



#### Identify houses within the buffer zones and create a categorical variable in the dataset indicating proximity to industrial zones
```{r}
  
# Convert the housing data to use the same CRS system
spdat2 <- spTransform( 
  spdat, # source polygon
  CRS( "+proj=longlat +datum=WGS84" ) # use the UTM projection, Syracuse is in Zone 18
  )
par( mar=c(0,0,0,0) )
plot( landUseNotInd, col = "grey", border = F )
plot( landUseInd, col = "orange", border = F, add = T )
plot( buffindzone, add=T, col=rgb(44,4,248,alpha=50, maxColorValue = 355))
#points( x = spdat2$lon, y = spdat2$lat, pch = 19 )
points( spdat2, cex = 50000 )
summary(spdat2[,1])
summary(landUseInd[,1])
summary(buffindzone[,1])

```

#### Create a buffer zone an eighth of a mile from schools, and plot the buffer zone.
```{r}

```

#### Identify houses within the buffer zones and create a categorical variable in the dataset indicating proximity to schools
```{r}

```

