---
title: "Lab 06 - Creating Buffers"
author: "Francisco Santamarina"
date: "March 09, 2017"
output:
  html_document:
    df_print: paged
    keep_md: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F )
```

## Part 1

Load the necessary packages.
```{r}
#library( censusAPI)
#library( ggmap )
#library( jsonlite )
library( maptools )
library( measurements )
#library( memisc )
#library( RCurl )
library( raster )
library( rgdal )
library( rgeos )
library( sp )
#library( spatialEco ) 

```

Load and clean the necessary dataset.
```{r}
# Set the working directory
#setwd( "//hd.ad.syr.edu/02/7796a1/Documents/Desktop/DDM/DDM II/Lab 06" )
setwd("~/Graduate School/PAI 690 Independent Study_DDM II/Labs/Lab 06_Buffers")
#dir.create( "shapefiles" )
setwd( "./shapefiles" )


# Download shapefiles
download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2010/TRACT/2010/tl_2010_36067_tract10.zip", "onondaga census tracts.zip" )
unzip( "onondaga census tracts.zip" )
file.remove( "onondaga census tracts.zip" )
dir()


# Load and plot shapefiles
syr <- readShapePoly( fn="tl_2010_36067_tract10", proj4string=CRS("+proj=longlat +datum=WGS84") )

par( mar=c(0,0,0,0) )
plot( syr,  border="gray10" )


# Add roads

download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2015/PRISECROADS/tl_2015_36_prisecroads.zip", "primary_and_secondary_roads.zip" )

unzip( "primary_and_secondary_roads.zip" )
file.remove( "primary_and_secondary_roads.zip" )
dir()

roads <- readShapeLines( fn="tl_2015_36_prisecroads", proj4string=CRS("+proj=longlat +datum=WGS84") )

head( roads@data )
table( roads$RTTYP )
class( roads )
#RTTYP = Road Type Code
# C = county
# I = interstate
# M = Common Name
# O = other
# S = state recognized
# U = U.S.
```

#### Create a polygon outline
```{r}
head( syr@data )
names( syr )
table( syr$GEOID10 )
str( syr$TRACTCE10 )

# Plot only the tracts within the city
# 0 or 1 in the 7th position of the FIPS ID string seem to differentiate city vs rural
# OR only looking at the last 4 is insufficient for a 3-digit census tract ID

last5tract <- unique(substr(syr$GEOID10, 7,11))
table(last5tract)
syr.sub <- syr[ (as.numeric(substr(syr$GEOID10, 7,11)) < 06200 ) & (as.numeric(substr(syr$GEOID10, 7,11)) > 0), ]
head(syr.sub)
plot(syr.sub)
class( syr.sub )
```


#### Clip Roads down to same size as census tracts
```{r}
# Using gClip() to do a modified subsetting of geospatial points/lines
# http://robinlovelace.net/r/2014/07/29/clipping-with-r.html
# rgeos version: 0.3-5, (SVN revision 447)
#  GEOS runtime version: 3.4.2-CAPI-1.8.2 r3921 
#  Polygon checking: TRUE

gClip <- function(shp, bb){  #shp = shapefile, bb = bounding box
  if(class(bb) == "matrix") b_poly <- as(extent(as.vector(t(bb))), "SpatialPolygons")
  else b_poly <- as(extent(bb), "SpatialPolygons")
  gIntersection(shp, b_poly, byid = T)
}

roads.clipped <- gClip(roads, syr)
#this returns a BOX that has the roads limited within it
plot( roads.clipped )
str( roads.clipped )
class( roads.clipped )
# Clip is computationally expensive, has to determine what to do with the polygons destroyed in the clip

# use intersect instead of gClip and syr.sub
## source: https://gis.stackexchange.com/questions/23931/how-to-intersect-lines-and-polygons-in-r
## author: xhie                library(raster)
## newLines <- raster::intersect(SpatialLinesDataframe, SpatialPolygonDataframe)
roads.cropped <- raster::intersect( roads, syr.sub )
plot( roads.cropped )
str( roads.cropped )
class( roads.cropped )
table( roads.cropped$RTTYP )

```


#### Extract the interstates as a separate layer
```{r}
#subset()

# Using roads.clipped

par( mar=c(0,0,0,0) )
plot( syr,  border="gray80" )
#plot( roads, col="steelblue", lwd=2, add=T )
plot( roads.clipped, col="steelblue", lwd=1, add=T )

#interstate <- roads[ roads$RTTYP == "I", ]
interstate1 <- roads.clipped[ roads.clipped$RTTYP == "I", ]
interstate.clipped <- gClip( interstate, syr )
#this returns a BOX that has the interstates limited within it
plot( interstate.clipped, col="red", add=T )



# Using roads.cropped

par( mar=c(0,0,0,0) )
plot( syr.sub,  border="gray80" )
#plot( roads, col="steelblue", lwd=2, add=T )
plot( roads.cropped, col="steelblue", lwd=1, add=T )

#interstate <- roads[ roads$RTTYP == "I", ]
interstate2 <- roads.cropped[ roads.cropped$RTTYP == "I", ]
interstate.cropped <- raster::intersect( interstate2, syr.sub )

plot( interstate.cropped, col="red", add=T )
class( interstate.cropped )
```

#### Create a buffer of approximately a quarter mile from the interstate
```{r}
#interstate.projected <- spTransform( interstate.cropped, CRS( "+init=epsg:3347" ) ) 
#coordinates(interstate.projected)

#class( interstate.projected )

# Approximate Buffer
buffhwy <- gBuffer(
  interstate.cropped,
  #interstate.projected,
  #width = 0.001, # width of buffer
  width = 0.005, # width of buffer
  byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
)
class( buffhwy )
# buffhwy <- SpatialPolygonsDataFrame( buffhwy, data=buffhwy@data)
class( buffhwy )
str( buffhwy )
plot( buffhwy, add=T)


# Accurate Buffer
# To make accurate buffers, use a variant of this code.
# Units: https://gis.stackexchange.com/questions/85300/units-of-gbuffer-from-rgeos-package
# Projection to use: https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/OverviewCoordinateReferenceSystems.pdf
interstate.projected.US <- spTransform( 
  interstate.cropped, # source polygon
  CRS( 
    "+proj=utm +zone=18 +datum=WGS84" ) # use the UTM projection, Syracuse is in Zone 18
  )

unit <- conv_unit( 10, "mi", "m" ) # converts from 0.25 mi to meters, the unit of UTM
buffhwy2 <- gBuffer(
    interstate.projected.US,
    #interstate.projected,
    #width = 0.001, # width of buffer
    width = unit,
    byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
  )
class( buffhwy2 )
# buffhwy <- SpatialPolygonsDataFrame( buffhwy, data=buffhwy@data)
str( buffhwy2 )
plot( buffhwy2, add=T)

```

#### Identify all houses that fall within the buffer zone
```{r}
dat <- read.csv( "https://raw.githubusercontent.com/R-Class/all-labs-ddmii-fjsantam/master/Lab%2006/data.csv" )
# For the code used to generate the raw CSV file above, refer to Lab_06_-_Geocoding_CSV_Script_FJS.Rmd in the Lab 06 folder on GitHub
names( dat )
nrow( dat )

# Convert the coordinates that are in lon (x) lat(y) order to spatial points, identifying the coordinates, data, and coordinate system
xy <- dat[ c( "lon", "lat") ]
spdat <- SpatialPointsDataFrame( xy, dat, proj4string=CRS("+proj=longlat +datum=WGS84")  )
class( spdat )
head( spdat )
points( spdat )
nrow( spdat )

# Function is: over( locations, layer )
homes.on.highway <- over( spdat, buffhwy )
class( homes.on.highway)

# buff1 is a SpatialPolygon; dat is a dataframe
# will return a list of points (top row) and which polygon they are in (represented by ID # or NA if not in one)
honh.logical <- homes.on.highway[is.na(homes.on.highway)]

honh.logical <- as.character(homes.on.highway) != "NA"
honh.logical[ is.na(honh.logical) ] <- FALSE
class( honh.logical )
```


#### Add a new categorical variable to the houses dataset: within buffer zone or not?
```{r}
dat <- as.data.frame( dat )
dat.buffer <- cbind( dat, honh.logical )
head( dat.buffer )
nrow( dat.buffer )

dat.hwy.buffer <- dat.buffer[ dat.buffer$honh.logical == TRUE, ]
nrow( dat.hwy.buffer )
head(dat.hwy.buffer)
```


#### Compile the Map
```{r}
par( mar=c(0,0,0,0) )
plot( syr.sub,  border="gray80" )
plot( roads.cropped, col="steelblue", lwd=1, add=T )
#plot( buffhwy, add=T)
plot( interstate.cropped, col="red", add=T )
points( spdat, cex = 0.5 )
points( x=dat.hwy.buffer$lon, y=dat.hwy.buffer$lat, pch=19, col="magenta" )
```


## Part 2

#### Create a buffer a quarter mile from industrial zones (LandUse) and plot the buffer zone
#### Redo this using the original map file and geocoding addresses in the parcels file, instead of the syr2 shapefile
There seem to be incompatibility issues with UTM and the other CRS system, given how they plot coordinates?
```{r}
#parcels <- read.csv( "https://raw.githubusercontent.com/lecy/maps-in-R/master/Data/syr_parcels.csv" )
#names(parcels)
#unique( parcels$LandUse )

### Get census tracts FIPS, grab all ids of 1-61, dissolve the polygons to get the outline
setwd( "~/Graduate School/PAI 690 Independent Study_DDM II/Labs/shapefiles" )
# download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2010/TRACT/2010/tl_2010_36067_tract10.zip", "onondaga census tracts.zip" )
# unzip( "onondaga census tracts.zip" )
# file.remove( "onondaga census tracts.zip" )
dir()
syr2 <- readShapePoly( fn="01-05-2015", 
         #proj4string=CRS("+proj=longlat +datum=WGS84") 
         proj4string=CRS( "+proj=utm +zone=18 +datum=WGS84" ) 
         )
head( syr2@data) #returns the output of a list of values in syr2 map data
class(syr2)

landUseInd <- syr2[ syr2$LandUse == "Industrial", ]
class(landUseInd)


# Approximate Buffer
buffindzone <- gBuffer(
  landUseInd,
  #width = 0.001, # width of buffer
  #width = 400, # width of buffer
  width = 2400, # expanding to play nice with the 189-house dataset
  byid = FALSE # applied across "subgeometries"   IF you do FALSE, will automatically create a dissolved buffer
)
class( buffindzone )

par( mar=c(0,0,0,0) )
plot( syr2,  border="gray80" )
plot( buffindzone, add=T)
```

#### Identify houses within the buffer zones and create a categorical variable in the dataset indicating proximity to industrial zones
```{r}
# Convert the housing data to use the same CRS system

spdat2 <- spTransform( 
  spdat, # source polygon
  CRS( "+proj=utm +zone=18 +datum=WGS84" ) # use the UTM projection, Syracuse is in Zone 18
  )
head(spdat2)
nrow(spdat2)

# Function is: over( locations, layer )
homes.in.ind <- over( spdat2, buffindzone )
class( homes.in.ind)

# buff1 is a SpatialPolygon; dat is a dataframe
# will return a list of points (top row) and which polygon they are in (represented by ID # or NA if not in one)
hini.logical <- homes.in.ind[is.na(homes.in.ind)]

hini.logical <- as.character(homes.in.ind) != "NA"
hini.logical[ is.na(hini.logical) ] <- FALSE
class( hini.logical )

#############
# Add to the dataset
dat.buffer <- cbind( dat.buffer, hini.logical )
head( dat.buffer )
nrow( dat.buffer )

dat.ind.buffer <- dat.buffer[ dat.buffer$hini.logical == TRUE, ]
nrow( dat.ind.buffer )
head(dat.ind.buffer)



plot( syr2,  border="gray80" )
plot( roads.cropped, col="steelblue", lwd=1, add=T )
#plot( buffhwy, add=T)
points( x=spdat2$lon, y=spdat2$lat, cex = 0.5 )
points( x=dat.ind.buffer$lon, y=dat.ind.buffer$lat, pch=19, col="magenta" )

```

#### Create a buffer zone an eighth of a mile from schools, and plot the buffer zone.
```{r}

```

#### Identify houses within the buffer zones and create a categorical variable in the dataset indicating proximity to schools
```{r}

```

